Some examples in extended logic counting classes

### 2015-07-21 tex/case2generalproof.tex
Correcciones menores y comentarios.

Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\Truc$. Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least $i\in\{1,\ldots,k\}$ such that there exists at least one assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_i(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$. Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$. As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
\begin{enumerate}
\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$. However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$. The assignment $\d$ to $\x$ satisfies that $\A\models\alpha_q(\d)$, so it holds that $\A\models\exists\x\:\varphi_q(\x,\P)$. It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$

@@

\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b). As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$. Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x)$, so $\A\models\psi_i(\P^\prime)$ for each $i$. But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
\end{enumerate}

@@

Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for any $i\in\{1,\ldots,k\}$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$. And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E1$

>>>>

Let $\varphi^\prime(\x,\X) = \varphi_1^\prime(\x,\X)\vee\cdots\vee\varphi_k^\prime(\x,\X).$ We will now show that $f_{\exists\x\varphi^\prime(\x,\X)} = f_{\exists\x\varphi(\x,\X)}\dotminus 1.$ Assume that $\A\in\Truc$. Suppose first that there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi(\x,\R).$ Let $q$ be the least index such that there exists an assignment $\R^\prime$ to $\X$ for which $\A\models\exists\x\:\varphi_q(\x,\R^\prime).$ Let $\d = (\d_1,\dots,\d_{n_q},\ep)$ be the lexicographically smallest assignment to $\x$ for which $\A\models\alpha_q(\d)$, where $\d_i$ is the corresponding assignment to $\x_i$, for every $i\in\{1,\dots,n_q\}$, and $\ep$ is an assignment for $\y$. Consider now the tuple $\P = (P_1,\dots,P_r)$ where $P_i = \bigcup_{j:\lambda_q(j)=i,j\in[1,n_q]}\{\d_j\}$. As we did in Lemma \ref{first} we will show that $\P$ is such that (a) $\A\models\exists\x\:\varphi(\x,\P),$ (b) $\A\not\models\exists\x\:\varphi^\prime(\x,\P),$ and (c) $\P$ is the only assignment to $\X$ that satisfies both (a) and (b)
\begin{enumerate}
\item[(a)] As we showed in part (a) of Lemma \ref{first}, if there is at least one assignment $\R$ to $\X$ such that $\A\models\exists\x\:\varphi_q(\x,\R)$, then $\A\models\exists\x\:\varphi_q(\x,\P)$ for this particular $\P$. However, as we showed in Claim \ref{alpha}, if there is an assignment $\s$ to $\x$ such that $\A\models\alpha_q(\s)$, then there {\em is} such an assignment to $\X$. The assignment $\d$ to $\x$ satisfies that $\A\models\alpha_q(\d)$, so it holds that $\A\models\exists\x\:\varphi_q(\x,\P)$. It immediately follows that $\A\models\exists\x\:\varphi(\x,\P).$

@@

\item[(c)] Suppose there is an assignment $\P^\prime \neq \P$ to $\X$ that satisfies both (a) and (b). As we deduce from the part (c) of Lemma \ref{first}, $\P$ is the only assignment to $\X$ such that $\A\not\models\:\psi_q(\P)$, so necessarily $\A\models\:\psi_q(\P^\prime)$. Since $\P^\prime$ assigned to $\X$ satisfies (a), then $\A\models\exists\x\:\varphi(\x,\P^\prime).$ By the choice of $q,$ every $i\in\{1,\ldots,q-1\}$ is such that $\A\models\forall\x\:\neg\alpha_i(\x)$, so $\A\models\psi_i(\P^\prime)$ for each $i$. But as we mentioned, also $\A\models\:\psi_q(\P^\prime),$ which means that $\A\models\exists\x\:\varphi^\prime_q(\x,\P^\prime),$ and so, $\A\models\exists\x\:\varphi^\prime(\x,\P^\prime),$ which leads to a contradiction.
\icristian{
Al igual que un argumento anterior, esta demostración es mas clara si es directa y NO por contradicción.	
}
\end{enumerate}

@@

Second, assume that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi_i(\x,\R)$ for every $i\in\{1,\ldots,k\}$. Let $\P$ be an arbitrary assignment to $\X$. Since $\A\not\models\exists\x\,\varphi_i(\x,\P)$, for any $i,$ we see that $\A\not\models\exists\x\,(\varphi(\x,\P)\wedge\chi(\x,\P))$ for any formula $\chi(\x,\P)$. It follows that there is no assignment $\R$ to $\X$ such that $\A\models\exists\x\,\varphi^\prime_i(\x,\R),$ for any $i\in\{1,\ldots,k\}$. And so, for every $\A\in\Truc$ such that $f_{\exists\x\,\varphi(\x,\X)}(\A)=0,$ we have that $f_{\exists\x\,\varphi^\prime(\x,\X)}(\A) = 0.$ Hence, from the results in the previous paragraph, if $f^\prime = f_{\exists\x\,\varphi^\prime(\x,\X)}$, we have that $f^\prime = f\dotminus 1\in\E1$.

\icristian{
Importante! Aqui falta argumentar porque la formula final esta en $\E1$.
}

### 2015-07-21 tex/th1.tex
Correcciones menores y comentarios.

%!TEX root = sharp-sigma-one-plus.tex
\begin{theo} \label{decisionptime}
The decision problem associated to a function in $\E1$ is in \textsc{P}.
\end{theo}
\begin{proof}
Let $f$ be a function in $\E1$. Then there is an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ such that
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
where $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$, $A = \{a_1,\ldots,a_{\vert A \vert}\}$, $\z$ is an $m$-tuple of variables and $\x$ is a $k$-tuple of variables. As said in \cite{DBLP:journals/jcss/SalujaST95}, a function $h:\Truc \to \N$ is in $\#\Sigma_0$ if there exists a quantifier-free $\L$-formula $\theta(\Y,\y)$ such that $h = f_{\theta(\Y,\y)}$. Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$. Moreover, let $g = f_{\psi(\X,\y)}$.
\begin{claim}
For each $\A \in \Truc$, $f(\A) > 0$ iff $g(\A) > 0$.
\end{claim}
\begin{proof}
($\Rightarrow$) Suppose $f(\A) > 0$. Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$. It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$. Therefore, $g(\A) > 0$. ($\Leftarrow$) Suppose $g(\A) > 0$. Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, be such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$. Therefore, $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
\end{proof}
Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula, so $g \in\Eo$. As we showed, for each $\A \in \Truc$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$, but as it is shown in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time. From this we conclude that the decision version of $f$ is in {\sc P}.
\end{proof}

>>>>

%!TEX root = sharp-sigma-one-plus.tex
\begin{theo} \label{decisionptime}
The decision problem associated to a function in $\E1$ is in \textsc{P}.
\end{theo}
\begin{proof}
Let $f$ be a function in $\E1$. Then there exists an extended quantifier-free $\L-$formula $\varphi(\x,\X,\z)$ where $\x$ is a $k$-tuple of variables and $\z$ is an $m$-tuple of variables such that
\begin{eqnarray*}
f(\A) &=& \mid \{ \langle\P,\e\rangle \mid \A \models \exists \x \, \varphi(\x,\P,\e) \} \mid,
\end{eqnarray*}
for every $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \Truc$ and $A = \{a_1,\ldots,a_{\vert A \vert}\}$. As said in \cite{DBLP:journals/jcss/SalujaST95}, a function $h:\Truc \to \N$ is in $\#\Sigma_0$ if there exists a quantifier-free $\L$-formula $\theta(\Y,\y)$ such that $h = f_{\theta(\Y,\y)}$. Let $\y = (\x,\z)$ and let $\psi(\X,\y) = \varphi(\x,\X,\z)$. Moreover, let $g = f_{\psi(\X,\y)}$.
\begin{claim}
For each $\A \in \Truc$, $f(\A) > 0$ iff $g(\A) > 0$.
\end{claim}
\begin{proof}
($\Rightarrow$) Suppose $f(\A) > 0$. Let $\P$ and $\e$ be such that $\A \models \exists \x \, \varphi(\x,\P,\e)$. It follows that there is at least one $\d \in A^k$ such that $\A \models \varphi(\d,\P,\e) = \psi(\P,(\d,\e))$. Therefore, $g(\A) > 0$. ($\Leftarrow$) Suppose $g(\A) > 0$. Let $\Q$ and $\c = (\c_1,\c_2)$, where $\c_1$ and $\c_2$ have $k$ and $m$ elements respectively, such that $\A \models \psi(\Q,\c) = \varphi(\c_1,\Q,\c_2)$. Therefore, $\A \models \exists \x \, \varphi(\x,\Q,\c_2)$, from which we conclude that $f(\A) > 0$.
\end{proof}
Note that $\psi(\X,\y)$ is an {\sc FO}-extended $\L$-formula and $g \in\Eo$. As we showed, for each $\A \in \Truc$, it holds that $\A \in L_f$ if and only if $g(\A) > 0$, but as it is shown in Theorem \ref{fp1}, $g(\A)$ can be evaluated in polynomial time. From this we conclude that the decision version of $f$ is in {\sc P}.
\end{proof}

### 2015-07-21 tex/th4.tex
Correcciones menores y comentarios.

\begin{theo}
$\E1$ is closed under substraction by one.
\end{theo}
\begin{proof}
Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$. Without loss of generality, let $\z = (z_1,\dots,z_d)$ and let $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}
\item \input{case1.tex}

>>>>

\begin{theo}
$\E1$ is closed under substraction by one.
\end{theo}
\begin{proof}
Consider an $\L$-formula of the form $\exists\x\:\varphi(\x,\X,\z)$ where $\z = (z_1,\dots,z_d)$ and $\X = (X_1,\dots,X_r)$. There are three possibilities regarding the size of the tuples of free variables $\z$ and $\X$: (1) $d>0$ and $r=0$ (2) $d=0$ and $r>0$ (3) $d,r>0$. This separates the proof in three cases:
\begin{enumerate}
\item \input{case1.tex}
